---
title: "CLT - Lecture"
author: "Luiz Bergo"
date: "03/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Central Limit Theorem - CLT

This is a brief review about the important Central Limit Theory commom called CLT. It is itent to show a practical approach, based om numeric and graphic presentation, so for further studies are highly recommended. 

The main idea is to apply the theory on a exponencial distribution and check whether is possible to conclude about the population distribution based on observed data.

The probability densidity function to an expondencial distribution takes form:

$f(x;\lambda)=\lambda e^{-x\lambda}$ , $x \ge 0$

For such distribution we have the population parameter as follow:

$\mu = \sigma = \frac {1}{\lambda}$

```{r exp_dist}

x <- seq(0,30, 0.01)
lambda = 0.2
dist_exp <- function(x,lambda) (lambda)*exp(-x*lambda)
y <- dist_exp(x, lambda)

plot(x,y, type= "l", col="red", main = "Exponencial probability density function ")
abline(v= 1/lambda, col= "blue" )
text(6, 0.15, "Theoretical mean", col = "blue", adj = c(0, -.1))

```

