---
title: "Tooth Growth Guinea Pig"
author: "Luiz Bergo"
date: "06/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
```

## Tooth Growth Guinea Pig

This document present analisys on dataset "ToothGrowth" available at R default packages. 

This is required as part of Statistical Inference course from Jhons Hopkins University powered by Coursera. 

The dataset is composed by 60 records of experiments on doses of *Vitamin C* administrated on a population of Guinea Pigs. The vitamin was introducted in pig´s diets in two forms as *Orange Juice* or *Ascorbic Acid* and the results on pig´s teeth group was recorded.

A summary of experiment table is shown:

```{r toothgowth}

data(ToothGrowth)
dt <- ToothGrowth

boxplot(dt$len~dt$supp, col=c("red","blue"), xaxt = "n", main = "Tooth growth for differente Vitamin C", xlab = "Type of Vitamin", ylab = "Tooth length (mm)")
axis(1, 1:2, c("Orange Juice", "Ascorbic Acid"))

```

Based on this graphic it is possible to observe that teeth wich pigs were fed with orange juice are bigger than that fed with scorbic acid.

Testing the confidence interval:

$H_{0}: \mu_{OC} = \mu_{VC}$
$H_{a}: \mu_{OC} != \mu_{VC}$

This is confirmed observing the fail to reject the hipotesys of the mean growth of two groups are the same.

Since it is not given any information about teeth grwoth population, a *t test* interval was evaluated considering both, equal and unequal variance between the two groups. 

```{r hipotesys_test}
t.test(OJ_dt$len,VC_dt$len, var.equal = FALSE , paired = FALSE)$conf
t.test(OJ_dt$len,VC_dt$len, var.equal = TRUE , paired = FALSE)$conf

```



In both cases the mean 

## Data investigation

Follow the summary main question is raised:

- Wich form of Vitamin C promote greater teeth growth?

Aiming for answer this question a simple box plot is presented.


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
