---
title: "Tooth Growth Guinea Pig"
author: "Luiz Bergo"
date: "06/11/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
```

## Tooth Growth Guinea Pig

This document present analisys on dataset "ToothGrowth" available at R default packages. 

This is required as part of Statistical Inference course from Johns Hopkins University powered by Coursera. 

The dataset is composed by 60 records of experiments on doses of *Vitamin C* administrated on a population of Guinea Pigs. The vitamin was introducted in pig´s diets in two forms as *Orange Juice* or *Ascorbic Acid* and the results on pig´s teeth group was recorded.

A summary of experiment table is shown:

```{r}

data(ToothGrowth)
dt <- ToothGrowth
str(dt)

```

Brief description of each column:

len: numeric value indicating growth of teeth.  
supp: the supplement type Orange Juice (OC) or Ascorbic Acid (VC).  
dose: diary dose in mg administred each day.  

## Data investigation

After observing the data a main question is raised:

- Wich form of *Vitamin C* promote greatest teeth growth?

Aiming answer this question a simple box plot is presented.


```{r toothgowth, fig.align='center', echo=FALSE}

boxplot(dt$len~dt$supp, col=c("red","blue"), xaxt = "n", main = "Tooth growth for different supplement", xlab = "Type of Vitamin", ylab = "Tooth length (mm)")
axis(1, 1:2, c("Orange Juice", "Ascorbic Acid"))

```

Based on this graphic it is possible to observe that teeth wich pigs were fed with orange juice are bigger than that fed with scorbic acid.

Testing the confidence interval:

$H_{0}: \mu_{OC} = \mu_{VC}  \\ H_{a}: \mu_{OC} \ne \mu_{VC}$ 


Since it is not given any information about teeth grwoth population, a *t test* interval was evaluated considering both, equal and unequal variance between the two groups. 


Critical value for a two sided teste with 95% confidence and 58 degrees of freedom:

```{r, echo = FALSE}

qt(0.975, 58)

```

Results for equal variances:

```{r hipotesys_test, echo = FALSE }

OJ_dt <- dt[dt$supp=="OJ",]
VC_dt <- dt[dt$supp=="VC",]
t.test(OJ_dt$len,VC_dt$len, var.equal = FALSE , paired = FALSE)

```

Results for different variances:

```{r, echo=FALSE}

t.test(OJ_dt$len,VC_dt$len, var.equal = TRUE , paired = FALSE)

```


The *alternative hyphotesis is reject* since in both cases the the confidence intervals include 0 and the observed t-value is inferior to the critical t-value. 

Although due to the narrow interval in the lower limit, it´s interesting evaluate the impact of doses in teeth growth.

```{r dose_supp, fig.align='center', echo=FALSE}
boxplot(dt$len~dt$supp*dt$dose, main = "Tooth growth for each supplement and dose",
        col=c("red","blue"), xaxt = "n", xlab = "Dose (mg/day)",ylab = "Tooth length ")
title(main = "Tooth growth for each supplement and dose")
axis(1, c(1.5, 3.5, 5.5 ) , c(0.5, 1.0, 2.0 ))
legend(x = c(4.5, 5.5), y = c(8, 12), legend=c("Orange Juice", "Ascorbic Acid"), 
       fill = c("red","blue"),col=c("red", "blue"), box.lty=0, cex=1.1)

```

By this graphic it is possible to see that the impact on growth are quite diferente for 0.5 and 1.0 mg/day. To reinforce this hypothesis tests comparing *Dose* and *Supp* for each interval are proposed.

Critical t-value for each dose are:

t-critical for dose of 0.5 mg/day with 95% of significance:

```{r, echo=FALSE}
OJ_dt05 <- OJ_dt[OJ_dt$dose==0.5,]
VC_dt05 <- VC_dt[VC_dt$dose==0.5,]
n05 <- length(OJ_dt05)+length(VC_dt05)
qt(0.975, n05)

```

t-critical for dose of 1.0 mg/day with 95% of significance:

```{r, echo=FALSE}
OJ_dt10 <- OJ_dt[OJ_dt$dose==1.0,]
VC_dt10 <- VC_dt[VC_dt$dose==1.0,]
n10 <- length(OJ_dt10)+length(VC_dt10)
qt(0.975, n10)
```

t-critical for dose of 2.0 mg/day with 95% of significance:

```{r, echo=FALSE}
OJ_dt20 <- OJ_dt[OJ_dt$dose==2.0,]
VC_dt20 <- VC_dt[VC_dt$dose==2.0,]
n20 <- length(OJ_dt20)+length(VC_dt20)
qt(0.975, n20)
```


*Testing the confidence interval for 0.5 mg/day:*

$H_{0}: \mu_{OC,dose=0.5} = \mu_{VC,dose=0.5}  \\ H_{a}: \mu_{OC,dose=0.5} \ne \mu_{VC,dose=0.5}$ 

```{r, echo=FALSE}

t.test(OJ_dt05$len,VC_dt05$len, var.equal = FALSE , paired = FALSE)

```

This result confirm the *sucess to reject the null hypothesis* for a dose of 0.5 mg/day.


*Testing the confidence interval for 1.0 mg/day:*

$H_{0}: \mu_{OC,dose=1.0} = \mu_{VC,dose=1.0}  \\ H_{a}: \mu_{OC,dose=1.0} \ne \mu_{VC,dose=1.0}$ 

```{r, echo=FALSE}

t.test(OJ_dt10$len,VC_dt10$len, var.equal = FALSE , paired = FALSE)

```

This result confirm the *sucess to reject the null hypothesis* for a dose of 1.0 mg/day.

*Testing the confidence interval for 2.0 mg/day:*

$H_{0}: \mu_{OC,dose=2.0} = \mu_{VC,dose=2.0}  \\ H_{a}: \mu_{OC,dose=2.0} \ne \mu_{VC,dose=2.0}$ 

```{r, echo=FALSE}

t.test(OJ_dt20$len,VC_dt20$len, var.equal = FALSE , paired = FALSE)

```
This result confirm the *fail to reject the null hypothesis* for a dose of 2.0 mg/day.

## Conclusion

The tooth growth are direct proporcional to dose of supplement. At 2.0 mg/day of supplement there is no evidence of differences in teeth growth. Although, *Orange Juice* at doses of 0.5 or 1.0 mg/day seems to intcrease tooth growth rates.